Figures for Paper

### 2020-11-26 paper_figures.Rmd
Update paper_figures.Rmd
Adding on the first VAF figures

```{r}
junc_colors = c("#648FFF","#fe6101")
ipsc_splicing = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/ipsc_splicing_results.csv")

@@

The 'C/G' tells which genotypes were supported by RNA-seq on rs12973192. 
The NB cell lines are het, as is the WTC11 cell line. SH-SY5Y cells are homozygote for the major allele. 
There was variability on the Klim hMN set on allelic expression. 
```{r}
rel_rna_cryptic_amount = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv")
rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3)]
rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
    ylab("UNC13A Cryptic PSI") + 
    xlab("") + 
    guides(color = FALSE)

@@

Only looking at disease relevant tissue and in patients
```{r}
detection_table = clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi), n = 2))) %>%
    mutate(stmn2_psi_groups = ifelse(is.na(stmn2_psi_groups),"  No STMN2",stmn2_psi_groups)) %>%
    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
                                        stmn2_psi_groups == 2 ~ "High STMN2",
                                        TRUE ~ stmn2_psi_groups)) %>%
  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
  group_by(call,unc13a_detected,stmn2_psi_groups) %>% 
  add_count(name = "genotype_detected") %>% 
  dplyr::select(call,unc13a_detected,genotype_detected,stmn2_psi_groups) %>% 
  unique() %>% 
  pivot_wider(names_from = "unc13a_detected",
              values_from = "genotype_detected") %>% 
  dplyr::rename(unc_not_detected = `FALSE`, unc_cryptic_detected = `TRUE`) %>%
  mutate(total_tissue = (unc_not_detected + unc_cryptic_detected)) %>% 
  mutate(detection_rate = unc_cryptic_detected / total_tissue) %>% 
  ungroup() %>% 
  mutate(stmn2_psi_groups = fct_relevel(stmn2_psi_groups, "No STMN2", after = 0)) %>% 
  mutate(detection_name = glue::glue("{call} \n ( {total_tissue} )"))

@@

>>>>

```{r}
junc_colors = c("#648FFF","#fe6101")
ipsc_splicing = fread(file.path(here(),"data","ipsc_splicing_results.csv")

@@

The 'C/G' tells which genotypes were supported by RNA-seq on rs12973192. 
The NB cell lines are het, as is the WTC11 cell line. SH-SY5Y cells are homozygote for the major allele. 
There was variability on the Klim hMN set on allelic expression. 
```{r}

rel_rna_cryptic_amount = fread(file.path(here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3)]
rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
    ylab("UNC13A Cryptic PSI") + 
    xlab("") + 
    guides(color = FALSE)

@@

Only looking at disease relevant tissue and in patients 
we see that the detection rate depends on both the level of TPD-43 pathology present, 
again I acknowledge STMN2 is a proxy, but it is what is is with the dataset, and the geneotype 
of the patients.

This difference is also not significant with a fisher's exact test, but
the

```{r}
detection_table = clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi), n = 2))) %>%
    mutate(stmn2_psi_groups = ifelse(is.na(stmn2_psi_groups),"  No STMN2",stmn2_psi_groups)) %>%
    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
                                        stmn2_psi_groups == 2 ~ "High STMN2",
                                        TRUE ~ stmn2_psi_groups)) %>%
  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
  group_by(call,unc13a_detected,stmn2_psi_groups) %>% 
  add_count(name = "genotype_detected") %>% 
  dplyr::select(call,unc13a_detected,genotype_detected,stmn2_psi_groups) %>% 
  unique() %>% 
  pivot_wider(names_from = "unc13a_detected",
              values_from = "genotype_detected") %>% 
  dplyr::rename(unc_not_detected = `FALSE`, unc_cryptic_detected = `TRUE`) %>%
  mutate(total_tissue = (unc_not_detected + unc_cryptic_detected)) %>% 
  mutate(detection_rate = unc_cryptic_detected / total_tissue) %>% 
  ungroup() %>% 
  mutate(stmn2_psi_groups = fct_relevel(stmn2_psi_groups, "No STMN2", after = 0)) %>% 
  mutate(detection_name = glue::glue("{call} \n ( {total_tissue} )"))

@@

## Variant allele expression in TDP-43 KD experiments

Across our KD's, we observed a clear allelic bias in the NB cells

```{r}
kd_vafs = fread(file.path(here(),"data","all_kds_unc13.snp.out.tsv"))

kd_vafs %>% 
    left_join(rel_rna_cryptic_amount) %>% 
    filter(!is.na(condition)) %>% 
    mutate(VAF = alt_reads / total) %>% 
    filter(condition != "control" ) %>% 
    filter(source %in% c("sh_dzap","shsy5y_normed","ipsc_normed_ward")) %>% 
    ggplot(aes(x = TARDBP,y = VAF, fill = source,size = total)) + 
    geom_point(pch = 21)  + 
    ylim(0,1)

kd_vafs %>% 
    left_join(rel_rna_cryptic_amount) %>% 
    filter(!is.na(condition)) %>% 
    filter(condition != "control" ) %>%     
    mutate(VAF = alt_reads / total) %>% 
    filter(source %in% c("sh_dzap","shsy5y_normed","ipsc_normed_ward")) %>% 
    dplyr::select(sample,ref_reads,alt_reads,TARDBP,source,VAF) %>% 
    data.table::melt(id.vars = c("sample","TARDBP",'source','VAF')) %>% 
    mutate(sample = fct_reorder(sample,-TARDBP)) %>% 
    ggplot(aes(x = sample,y = value, fill = variable)) + 
    geom_col(pch = 21, size = 4) +
    coord_flip()
```

Also look at the coverage in the Liu nuclear facs sorted neurons

```{r}
liu_vafs = fread(file.path(here(),"data","liu_facs_neurons_unc13.snp.out.tsv"))
liu_vafs %>% 
    mutate(VAF = alt_reads / total) %>% 
    data.table::melt(id.vars = c("sample",'VAF','total')) %>% 
    mutate(allele = ifelse(variable == "ref_reads", "C","G")) %>% 
    mutate(sample_name = sub("_TDP_.*", "", sample)) %>% 
    mutate(sample_name = gsub("_unsorted", "", sample_name)) %>% 
    mutate(condition = sub(".*_", "", sample)) %>% 
    ggplot(aes(x = condition,y = value, fill = allele)) + 
    geom_col(pch = 21, size = 4) +
    facet_wrap(~sample_name,scales = "free_y")
```
